\guidelinesubsubsection{Rationale}

A \emph{session} is any bounded stretch of activity during which an LLM is invoked. Sessions cover one-shot prompts, batch runs, multi-turn conversations, and agentic runs that span many tool calls.
A \emph{session trace} records everything that crosses the LLM boundary or is produced around the model during that stretch: prompts received, responses returned, tool calls the model made together with their arguments and results, plans the model developed, and which of the statically configured artifacts (e.g., skills, context files, sub-agents, tools) were actually picked up at runtime.

Two kinds of session trace are important to capture.
An \emph{interaction log} captures what humans or other software tools sent to the LLM and what the LLM returned.
A \emph{runtime trace} captures what the LLM did at runtime: calls to external tools (e.g., APIs, file systems, databases, MCP servers, sub-agents) and activations of configured artifacts (e.g., context files, skills). Reporting both kinds is needed for an agentic run, because neither tells the full story on its own.

Even with the exact same prompts, decoding strategies, and parameters, LLMs can behave non-deterministically.
Non-determinism can arise from probabilistic sampling and, even with greedy decoding (temperature = 0), from batching and floating-point arithmetic on GPUs~\cite{YuanNondeterminismLLMInference}, and from Mixture-of-Experts routing~\cite{Chann2023}.
For other researchers to verify the conclusions drawn from LLM interactions, the traces of what happened at runtime are often as important as the design of the system itself.
This matters particularly for studies targeting commercial software-as-a-service (SaaS) solutions such as ChatGPT, and for agentic runs where behavior depends on how the model chose among many possible tool calls.

The rationale is similar to reporting interview transcripts in qualitative research.
Just as a human participant might give different answers to the same question asked two months apart, the responses from tools such as ChatGPT can also vary over time, and the trace of an agent's decisions on any given run is often not reproducible at a later date.
This guideline addresses runtime reporting and complements \design, which covers the static artifacts that determine what the model sees.

\guidelinesubsubsection{Recommendations}

\paragraph{Interaction Logs:}
Researchers \should report the full interaction logs (prompts sent to the LLM and responses returned) as part of their \supplementarymaterial.
For agentic systems, interaction logs generalize to cover exchanges between humans and the agent as well as between external tools and the agent, including human-in-the-loop feedback, approval or rejection decisions, and iterative refinements.
These \should also be reported as \supplementarymaterial so that readers can reconstruct the sequence of exchanges and assess human oversight decisions.
For traces containing sensitive information, researchers \must anonymize personal identifiers, replace proprietary code with placeholders, and clearly highlight modified sections.

\paragraph{Runtime Traces:}
When an LLM calls out to external tools (e.g., APIs, file systems, databases, MCP servers, sub-agents) or activates configured artifacts (e.g., context files, skills, sub-agents reported under \design), this runtime activity forms a \emph{runtime trace} distinct from the interaction log.
Researchers \should report the complete runtime trace as \supplementarymaterial, including for each entry the tool or artifact name, arguments (if any), result, and ordering relative to surrounding interaction-log entries.
This lets readers attribute task success to the model, the external tools, or their interaction pattern, and distinguish artifacts that were configured from those that actually influenced a given run.
Researchers \should record runtime traces in an open format compatible with the OpenTelemetry Protocol (OTLP), such as the OpenTelemetry GenAI semantic conventions~\cite{otel_genai_semconv} or OpenInference~\cite{openinference_spec}, and report the version used, rather than a study-specific schema.

\paragraph{Agentic Plans:}
For agentic systems that autonomously plan and execute tasks, the developed plans \should be reported as \supplementarymaterial if available.
In Claude Code, for example, a plan is a short Markdown document the user can open and edit during a session, listing the proposed steps and the files or commands the agent intends to touch.
Other frameworks such as LangGraph keep plans inside the agent's internal execution state.

All reported traces \must be made publicly available as \supplementarymaterial, subject to privacy and confidentiality constraints. When full trace logging is not feasible, researchers \should provide representative examples or anonymized traces.

\guidelinesubsubsection{Examples}

An example of reporting full interaction logs is the study by \citeauthor{ronanki2023investigating}~\cite{ronanki2023investigating}, for which the authors reported the full answers of ChatGPT and uploaded them to \href{https://zenodo.org/records/8124936}{Zenodo}.
For agentic systems, \citeauthor{DBLP:conf/kbse/BouzeniaP25}~\cite{DBLP:conf/kbse/BouzeniaP25} unified the runtime trajectories of three SE agents (\emph{RepairAgent}, \emph{AutoCodeRover}, \emph{OpenHands}) into a common format and released the resulting 120 trajectories with 2,822 LLM interactions as a public dataset.

\guidelinesubsubsection{Benefits}

Unlike human participant conversations, which often cannot be reported because of confidentiality, LLM interaction logs can be shared.
This enables reproduction studies, tracking of response changes over time or across model versions, and secondary research on LLM consistency for specific SE tasks.

For agentic systems, reporting runtime traces alongside interaction logs lets readers attribute task outcomes to the right component: the model, the tool, or the orchestration pattern.
Usage traces complement the static configuration reported under \design by showing which of the configured artifacts actually influenced a given run.

\guidelinesubsubsection{Challenges}

Not all systems allow reporting of complete interaction logs with ease, and this hinders transparency and verifiability.
For commercial tools, researchers \must report all available information and acknowledge unknown aspects as limitations.
Tool-call traces for commercial SaaS agents are often opaque: the user sees the final response but not the sequence of internal tool calls.
When this is the case, authors should document what was and was not observable.
For agents running locally or via open-source tools, these traces are usually more accessible and \should be reported whenever available.
Agent frameworks differ in whether and how they log agent-to-agent communication, so reporting practices vary across studies.

\guidelinesubsubsection{Study Types}

Runtime trace reporting requirements depend heavily on the study type and on the accessibility of the underlying system.
The general expectation is that interaction logs are reported as \supplementarymaterial whenever feasible; runtime traces are additionally expected when agentic execution or evaluation harnesses go beyond direct API calls.

For \llmusage, especially observational studies targeting commercial tools, researchers \must report the full interaction logs except when transcripts might identify anonymous participants or reveal personal or confidential information.
If complete interaction logs cannot be shared (e.g., because they contain confidential information), the prompts and responses \must at least be summarized and described in the \paper.
For \newtools that use agentic execution, researchers \should report runtime traces for the runs used to evaluate the tool.
For \benchmarkingtasks that use agent-based harnesses, researchers \should report runtime traces for representative runs, letting readers understand how task success depends on the agent's decision sequence rather than the model's raw output alone.
For \annotators, \judges, \synthesis, and \subjects, when the research setup involves multi-turn interaction or agentic orchestration, researchers \should report the corresponding interaction logs and, where applicable, runtime traces.

\guidelinesubsubsection{Advice for Reviewers}

As with other guidelines, missing trace information is typically a minor revision request unless so much is missing that methodological rigor cannot be assessed.
Reviewers should recognize that complete trace reporting is easier for local and open-source setups than for commercial SaaS agents.
When reviewing commercial-tool studies, reviewers should focus on whether authors have reported everything the system exposed and have been explicit about what remains unobservable.

\guidelinesubsubsection{See Also}
\begin{itemize}[label=$\rightarrow$]
    \item \seedesign: Session traces show runtime behavior; system and prompt design covers the static artifacts that produce it.
    \item \seemodelversion: A trace cannot be reproduced or compared across studies without the model identity that produced it.
    \item \seehumanvalidation: Agentic traces produce outputs that often warrant human evaluation.
    \item \seeopenllm: Open models let other researchers reproduce a reported trace on the exact same model weights.
    \item \seelimitationsmitigations: When a tool hides parts of the trace, authors must report the gap as a limitation.
\end{itemize}
