\guidelinesubsubsection{Rationale}

Reproducibility depends on access to the model under study.
When research relies exclusively on proprietary models, other researchers cannot independently verify or build upon the findings.
Including an open LLM as a baseline ensures that at least part of the study can be fully replicated.

\guidelinesubsubsection{Recommendations}

Empirical studies using LLMs in SE, especially those that target commercial tools or models, \should incorporate an open LLM as a baseline and report established metrics for inter-model agreement.
We acknowledge that including an open LLM baseline might not always be possible, for example, if the study involves human participants, and letting them work on the tasks using two different models might not be feasible.
Using an open model as a baseline is also not necessary if the use of the LLM is tangential to the study goal.

Open models allow other researchers to verify research results and build upon them, even without access to commercial models.
A comparison of commercial and open models also allows researchers to contextualize model performance.
Researchers \should provide a complete replication package as part of their \supplementarymaterial, including clear step-by-step instructions on how to verify and reproduce the results reported in the paper.

Open LLMs are available on platforms such as \href{https://huggingface.co/}{\emph{Hugging Face}} and can be hosted locally using frameworks such as \href{https://ollama.com/}{\emph{Ollama}} or \href{https://lmstudio.ai/}{\emph{LM Studio}}, or on cloud services such as \href{https://together.ai/}{\emph{Together AI}}, AWS, Azure, and Google Cloud.

The term ``open'' can have different meanings in the context of LLMs.
\citeauthor{widder2024open} discuss three types of openness (transparency, reusability, and extensibility) and what openness can and cannot provide~\cite{widder2024open}.
The \emph{Open Source Initiative} (OSI)~\cite{OSIAI2024} defines open-source AI as having access to everything needed to understand, modify, share, retrain, and recreate the model.

Researchers can also explore open-source tools such as Continue, Cline, and opencode as alternatives to commercial tools like GitHub Copilot and Claude Code, enabling instrumentation, architectural transparency, and detailed telemetry collection.

\guidelinesubsubsection{Examples}

An increasing number of studies have adopted open LLMs as baseline models.
For example, \citeauthor{wang2024and} evaluated seven advanced LLMs, six of which were open-source, testing 145 API mappings drawn from eight popular Python libraries across 28,125 completion prompts aimed at detecting deprecated API usage in code completion~\cite{wang2024and}.
\citeauthor{moumoula2024large} compared four LLMs on a cross-language code clone detection task~\cite{moumoula2024large}.
Three evaluated models were open-source.
\citeauthor{gonccalves2025evaluating} fine-tuned the open LLM \emph{LLaMA} 3.2 on a refined version of the \emph{DiverseVul} dataset to benchmark vulnerability detection performance~\cite{gonccalves2025evaluating}.
\emph{CodeBERT}, a bimodal transformer pre-trained by Microsoft Research, has been widely used as an open baseline~\cite{DBLP:journals/jss/YangZCZHC23, DBLP:conf/gaiis/XiaSD24, DBLP:conf/kbse/SonnekalbGBM22, DBLP:conf/icse/CaiYMMN24}, with model weights, source code, and data-processing scripts published on GitHub~\cite{codebert}.

\guidelinesubsubsection{Benefits}

A true open LLM baseline improves reproducibility by exposing model architectures, parameter settings, and ideally training data, enabling independent verification of results.
Moreover, by adopting an open-source baseline, researchers can directly compare novel methods against established performance metrics without the variability introduced by proprietary systems.
These models allow detailed inspection of data processing pipelines and decision-making routines, helping identify biases and model limitations.
Furthermore, unlike closed-source alternatives, which can be withdrawn or altered without notice, open LLMs ensure long-term accessibility and stability, preserving critical resources for future studies.
Finally, the licensing requirements associated with open-source implementations lower financial barriers, making advanced language models attainable for research groups operating under constrained budgets.

\guidelinesubsubsection{Challenges}

Open-source LLMs face several notable challenges.
First, they often lag behind the most advanced proprietary ``frontier'' models in common benchmarks, making it difficult for researchers to demonstrate clear improvements when evaluating new methods using open LLMs alone.
Additionally, deploying and experimenting with these models typically requires substantial hardware resources, in particular high-performance GPUs, which may be beyond reach for many academic groups.
The notion of ``openness'' remains in flux: many models release only trained weights without training data or methodological details (``open weight'' openness)~\cite{Gibney2024}, which is why we reference the OSI definition in our recommendations~\cite{OSIAI2024}.
Finally, unlike APIs provided by proprietary vendors (e.g., the OpenAI API), installing, configuring, and fine-tuning open-source models can be technically demanding.
Documentation is often sparse or fragmented, placing a high barrier to entry for researchers without specialized engineering support.

\guidelinesubsubsection{Study Types}

This guideline applies primarily to study types in which the researcher controls which LLM is used.
In formal benchmarking studies and controlled experiments (see \benchmarkingtasks), an open LLM \must be one of the models under evaluation.
When evaluating \newtools, researchers \should use an open LLM as a baseline whenever it is technically feasible; if integration proves too complex, they \should report the initial benchmarking results of open models.
For \annotators and \judges, researchers \should compare annotation or judgment quality from open vs.\ commercial models to assess the extent to which results depend on a specific proprietary model.
For \synthesis, researchers \should compare synthesis results from open and commercial models to evaluate robustness of the findings.
For \llmusage, using an open LLM as a baseline is often not feasible when the study observes participants using specific commercial tools; in such cases, investigators \should explicitly acknowledge its absence and discuss how this limitation might affect their conclusions.
For \subjects, using an open LLM as a baseline may similarly be impractical if the study design requires the capabilities of a specific model; researchers \should acknowledge this limitation when applicable.

\guidelinesubsubsection{Advice for Reviewers}

Reviewers should distinguish between LLM use that is central to the research (e.g., building LLM-driven SE tools, using LLMs for synthesis) and use that is tangential (e.g., generating recruiting materials). An open LLM baseline is expected only when LLM use is central; otherwise, its absence need not be justified.
When an open baseline is expected, reviewers should look for either the use of an open LLM or a convincing argument for why it is impractical. If authors claim openness, some justification for that characterization is appropriate. Where practical, reviewers should examine whether the replication package contains sufficiently detailed instructions.
Reviewers should \textit{not} penalize studies for performance differences between open and proprietary models, as this is beyond the authors' control.

\guidelinesubsubsection{See Also}
\begin{itemize}[label=$\rightarrow$]
    \item \seemodelversion: Authors must report version, configuration, and parameters for open models, not just commercial ones.
    \item \seedesign: Self-hosting an open model adds hardware, hosting, and infrastructure to document.
    \item \seebenchmarksmetrics: Inter-model agreement metrics make open-vs-commercial comparisons interpretable.
    \item \seelimitationsmitigations: When using an open LLM as a baseline is impractical, authors must report this as a limitation.
\end{itemize}
