\guidelinesubsubsection{Rationale}

Even with well-justified benchmarks and metrics (\benchmarksmetrics), automated measurement captures only the constructs the benchmark was designed to measure.
Subjective constructs and constructs that no current benchmark covers require human assessment.

\guidelinesubsubsection{Recommendations}

When LLMs automate research or software development tasks previously performed by humans, the LLM's performance needs to be assessed.
Where no benchmark adequately operationalizes the target construct, researchers \should validate LLM outputs against human judgment.

\paragraph{Study Design Considerations:}
Studies that include human participants need additional considerations, including a recruitment strategy, annotation guidelines, training sessions, or ethical approvals.
Therefore, researchers \should consider human validation early in the study design, not as an afterthought.
Authors \must clearly define in the \paper the constructs that the human and LLM annotators evaluate~\cite{DBLP:conf/ease/RalphT18}.
When designing custom instruments to assess LLM output (e.g., questionnaires, scales), researchers \must share their instruments.

\paragraph{Subjective Judgment and Agreement:}
When the judgment is subjective (i.e., depends on the judge's values or theories), the LLM output \should be evaluated against judgments aggregated from multiple human judges, and researchers \should clearly describe their aggregation method and reasoning. One systematic approach is to randomly order the objects requiring judgment and put them in groups small enough to judge in one to three hours. Two or three human experts review the first group together, simultaneously judging, discussing, and creating a set of decision rules documenting their reasoning. Then, the experts iterate among rounds of:
\begin{enumerate}
    \item independently rating one group of objects,
    \item calculating inter-rater agreement (IRA) or reliability (IRR),
    \item meeting to discuss disagreements and reach consensus,
    \item updating the decision rules so the disagreement will not recur.
\end{enumerate}

The goal is to reach a target IRA or IRR (e.g., Krippendorff's $\alpha>0.8$), indicating sufficient decision rules. Once this target is reached and sustained for two to three groups, it is permissible to continue with a single rater.
Researchers \should report measures of IRA or IRR, preferably broken down by round.

Confounding factors \should be discussed and, where feasible, controlled for (e.g., by categorizing participants according to their level of experience or expertise).
For value-laden or culturally contingent constructs (e.g., judging code-style appropriateness or comment helpfulness), researchers \should describe rater demographics beyond expertise (e.g., geographic, linguistic, or professional background) and discuss potential demographic biases in rater recruitment and instructions~\cite{bean2025measuring}.
Where applicable, researchers may perform a power analysis~\cite{Cohen1992,DBLP:journals/infsof/DybaKS06} to estimate the required sample size, ensuring sufficient statistical power in their experimental design.
However, we also note that to our knowledge, there is limited established guidance specific to determining sample sizes for LLM-human comparisons. In other fields, 100 comparisons are often made without further explanation~\cite{DBLP:journals/npjdm/TamSKSPMOWVFMCSPW24}.

Researchers \should use established reference models to compare humans with LLMs.
For example, \citeauthor{DBLP:conf/icse/SchneiderFW25}~\cite{DBLP:conf/icse/SchneiderFW25} outline design considerations for studies comparing LLMs with humans.
If studies aim to automate annotating software artifacts using an LLM, researchers \should follow systematic approaches to decide whether and how human annotators can be replaced (e.g., showing that a jury of three LLMs exhibit model-to-model agreement comparable to trained human annotators, such as Krippendorff's $\alpha>0.8$). Model-to-model agreement requires the same threshold as human-to-human agreement, since low thresholds may simply reflect shared model biases rather than annotation quality.

\paragraph{Agentic Tools:}
Besides generating and modifying content, agentic software development tools such as \emph{Claude Code} can autonomously call command-line tools or pull in additional information from MCP servers.
For such tools, it makes sense to separate textual output (e.g., file changes) from output related to tool execution.
When evaluating agentic tools, researchers \should assess the feedback that users provided for proposed changes, report statistics on how frequently they accepted the content, and how they modified it.
Agentic human-in-the-loop interaction is a built-in human validation, with larger degrees of freedom than traditional experiments that use LLMs directly.

\guidelinesubsubsection{Examples}

\citeauthor{DBLP:conf/msr/AhmedDTP25}~\cite{DBLP:conf/msr/AhmedDTP25} proposed a systematic method for deciding whether LLMs can replace human annotators on a given task, using model-to-model agreement as an initial screening criterion (with a threshold of $\alpha>0.5$) and model confidence for sample-level decisions. However, their threshold is well below the levels generally considered acceptable for inter-rater agreement.
\citeauthor{Krippendorff2018} recommends discarding data with $\alpha<0.667$, considers $0.667 \leq \alpha < 0.8$ sufficient only for tentative conclusions, and requires $\alpha \geq 0.8$ for reliable data~\cite{Krippendorff2018}. Researchers adopting similar screening approaches should apply thresholds consistent with these established standards. Notably, while three LLMs exhibited higher inter-model agreement than human annotators on some tasks, human-model agreement remained low on others.
This illustrates that high model-model reliability does not guarantee alignment with human judgment, reinforcing the need for human validation before assuming that LLM annotations are valid.
Moreover, a high model-model agreement could merely indicate that the models share systematic biases and hence reliably agree on the wrong answer.

\citet{DBLP:journals/corr/abs-2501-19297}~\cite{DBLP:journals/corr/abs-2501-19297} evaluated ChatGPT's capability to generate requirements documents by comparing an LLM-generated and a human-generated document based on the same business use case. Domain experts reviewed both documents and attempted to distinguish their origin.

\guidelinesubsubsection{Benefits}

Validation against human judgments builds confidence in the LLM's accuracy and validity, increasing the trustworthiness of findings, especially when IRA or IRR metrics are reported~\cite{khraisha2024canlargelanguagemodelshumans}. Furthermore, incorporating feedback from human judges may help to improve the LLM or LLM-based tool based on the reported experiences.

\guidelinesubsubsection{Challenges}

Assessment of a construct such as LLM performance is challenging because ensuring that a construct is defined well and operationalized using an appropriate measurement model requires a deep understanding of (1) the construct, (2) construct validity in general, and (3) instrumentation~\cite{DBLP:journals/tse/SjobergB23,DBLP:books/sp/24/RalphKAA24}. Comparing an LLM to human judges is typically slower and more expensive than machine-generated measures. More fundamentally, neither human judgment nor machine-generated measures provides an objective ground truth against which LLM accuracy can be firmly determined.   

Human judgments exhibit variability due to differences in experience, expertise, interpretations, and personal biases~\cite{DBLP:journals/pacmhci/McDonaldSF19}. When diverse humans rate items reliably given clear decision rules, we assume that reliability implies validity, but it does not. Measuring reliability is \textit{much} easier than measuring validity, and often the best researchers can do is argue conceptually for why their judges, decision rules, and constructs \textit{should} produce valid ratings.

Researchers must weigh the cost and time need for human validation against the expected corresponding improvement in validity over machine-generated measures. 

\guidelinesubsubsection{Study Types}

This guideline applies to all study types, although the need for human validation varies.
For \annotators, researchers \should validate LLM-generated annotations against human annotators to assess labeling quality and identify systematic biases.
When using \judges, researchers \should co-create initial rating criteria with humans and validate a sample of LLM judgments against human expert assessments.
For \synthesis, researchers \should employ human oversight to verify that qualitative interpretations and synthesized outputs faithfully represent the underlying data.
For \subjects, researchers \should validate simulated responses against real human data to assess the fidelity of the simulation.
For \llmusage, researchers \should carefully reflect on the validity of their evaluation criteria and validate subjective assessments with human experts.
For \newtools whose output is supposed to match human expectations, researchers \should validate the LLM output against human judgment.
For \benchmarkingtasks, there is less need for human validation when using extensively validated and widely-used benchmarks, but researchers \should employ human validation when creating or adapting new benchmarks.

\guidelinesubsubsection{Advice for Reviewers}

Human validation may be the most challenging of our guidelines to assess because it often requires evaluating conceptual arguments. If LLM output is validated only by comparison with other LLMs, reviewers should look for \textit{quantitative empirical evidence} that such comparison is reliable \textit{and} valid. High inter-model agreement alone is insufficient, as reliability does not imply validity. Similarly, reviewers should expect evidence that any employed benchmarks are reliable and valid. Absent such evidence, human validation is warranted.
A single human judge is appropriate only when judgments depend on widely accepted theories and involve limited value conflict (e.g., tagging method names containing abbreviations). For multiple judges, reviewers should expect IRA/IRR improvement techniques as described in the recommendations above (experienced raters, organized rounds, consensus meetings, updated decision rules). Low IRA or IRR (e.g., Krippendorff's $\alpha<0.8$) without these techniques is a concern. Conversely, if authors have followed best practices and still obtained mediocre results (e.g., $0.66<\alpha<0.8$), this should be noted as a limitation.
Beyond reliability, reviewers should expect authors to explain conceptually why their human judgments should be valid, considering construct definitions, decision rules, and judge expertise.
As with other guidelines, missing information is typically a revision request. Absent judge instructions, instruments, decision rules, or construct definitions may prevent assessment of rigor and validity, while missing recruitment details are less critical. Clarification requests about construct definitions are routine and should not alone warrant rejection.