Researchers \must describe the full architecture of LLM-based tools that they develop in their \paper, including the role of the LLM, interactions with other components, and the overall system behavior. If autonomous agents are used, researchers \must specify agent roles, reasoning frameworks, and communication flows. Hosting, hardware setup, and latency implications \must be reported. For tools using retrieval or augmentation methods, researchers \must describe how external data is retrieved, stored, and integrated; data preprocessing, versioning, and update frequency \should also be reported. For ensemble architectures, the coordination logic between models \should be explained in the \paper. Researchers \should justify design decisions. Where legally possible, researchers \should release the source code of their implementation under an open-source license. Non-disclosed confidential or proprietary components \must be acknowledged as reproducibility limitations. Researchers \must also describe any context-file mechanisms used (e.g., \texttt{AGENTS.md}, \texttt{CLAUDE.md}) and \should publish the file contents as \supplementarymaterial. Researchers \must summarize in the \paper which tools and skills were exposed to the model, and \should include the complete list (names with one-line purposes), tool schemas, skill definitions, sub-agent definitions, and connected MCP servers as \supplementarymaterial. Researchers \must publish all prompts or, when using templates, prompt templates with representative instances, including their structure, content, formatting, and dynamic components, as \supplementarymaterial. If full prompt disclosure is not feasible, for example because of privacy or confidentiality concerns, summaries or examples \should be provided. The prompting strategy used (e.g., zero-shot, one-shot, few-shot) \must be specified, and the development rationale and selection process \should be described. When prompts are long or complex, input handling and token optimization strategies \should be documented. For dynamically generated prompts, the generation process \must be documented; for user-authored prompts, researchers \should describe how they were collected and analyzed. Prompt reuse across models and configurations \must be specified. Researchers \should report prompt evolution and any LLM-suggested refinements.
