\scopeparagraph{Software Engineering as our Target Discipline}

We target SE research because reporting guidelines from other disciplines do not address its specific needs (see Related Reporting Guidelines below). SE research employs a wide variety of empirical methods~\cite{ralph2021empiricalstandardssoftwareengineering, DBLP:books/sp/WohlinRHORW24}. We organize the LLM-involving subset of these methods into a taxonomy of \studytypes\ifpaper~(Section~\ref{sec:study-types})\fi, and each of our \guidelines specifies its applicability per study type.

\scopeparagraph{Related Reporting Guidelines}

Reporting guidelines have a long tradition in healthcare research. CONSORT~\cite{Schulz2010}%
% GROUNDING: Schulz, Altman, and Moher. ``CONSORT 2010 Statement: updated guidelines for reporting parallel group randomised trials.'' BMJ 340:c332. The CONSORT 25-item checklist is the canonical structural template for randomized clinical trial reports since 2010; the CONSORT extension family includes CONSORT-AI for AI clinical trials. CONSORT and TRIPOD are separate healthcare reporting traditions: CONSORT covers randomized trials, TRIPOD covers prediction-model studies.
{} for randomized clinical trials (since 2010) is one canonical example, and our reporting checklist (\ifpaper Appendix~\ref{sec:checklist} \else \href{/checklist/}{checklist} \fi) follows its structural template. \citeauthor{Gallifant2025}~\cite{Gallifant2025}'s TRIPOD-LLM is a recent reporting guideline for LLM studies in healthcare%
% GROUNDING: Section 1 of Gallifant et al.: ``Here, we report the TRIPOD-LLM statement, an extension of TRIPOD+AI, developed to address these unmet needs.'' The TRIPOD-LLM 19-item main checklist (verified against the PMC version) includes items on the clinical use case (item 3a Background/Context, item 3b Target Population), public involvement (item 15), and a Discussion subsection on usability context, intended use, input handling, and user expertise (items 19c-19f); no item addresses tool architecture, software-system engineering, or SE-specific evaluation metrics. TRIPOD-LLM does not cite CONSORT; it extends the TRIPOD lineage rather than the CONSORT lineage.
{}. TRIPOD-LLM's checklist includes items specific to healthcare, such as clinical care usability, and does not address tool architectures, benchmarking, or code-specific evaluation that SE research requires. Other communities have introduced LLM-specific reporting guidelines more recently. \citeauthor{DBLP:conf/chi/NavarroSA26}~\cite{DBLP:conf/chi/NavarroSA26}%
% GROUNDING: Section 2: ``This paper focuses on LLM-integrated systems reported in HCI papers, defined here as systems that query an LLM at any stage of program execution.''
{} address HCI papers reporting LLM-integrated systems, defined by the authors as \enq{systems that query an LLM at any stage of program execution}, and \citeauthor{Kapoor2024REFORMS}~\cite{Kapoor2024REFORMS}%
% GROUNDING: \emph{Science Advances} 10(18), 2024: a 32-item field-agnostic checklist for machine-learning-based science, organized into eight sections (study goals, computational reproducibility, data quality and provenance, data preprocessing, modeling, data leakage, metrics and uncertainty, generalizability and limitations).
{} propose REFORMS, a field-agnostic checklist for machine-learning-based science. In SE itself the tradition is younger. Our work builds on \citeauthor{DBLP:conf/icse/SallouDP24}'s vision paper on mitigating validity threats in LLM-based SE research~\cite{DBLP:conf/icse/SallouDP24}%
% GROUNDING: Sallou, Durieux, Panichella. ``Breaking the Silence: the Threats of Using LLMs in Software Engineering'' (ICSE NIER 2024). Catalogs three categories of LLM-specific threats to validity (closed-source models, implicit data leakage, reproducibility) and pairs each with mitigation strategies (versioned model archives, metamorphic test data, multiple replication runs with variability metrics, detailed execution metadata).
{} and our own earlier workshop position paper~\cite{DBLP:conf/wsese/0001BFB25}, providing more detailed advice for a wider variety of research methods. \citeauthor{DBLP:journals/corr/abs-2601-01954}~\cite{DBLP:journals/corr/abs-2601-01954}%
% GROUNDING: Section 1: ``We analyzed nearly 300 papers published in ICSE, FSE, and ASE since 2022 [...] We surveyed 105 Program Committee (PC) members of the aforementioned conferences.'' Section 5 (Table 5 and Section 5.2 Discussion of Practices) reports the exact LLM version in only 16.43\% of papers (E1), prompt construction justification in 58.74\% (E5), and prompting as a threat to validity in 23.43\% (E9); PC member essentiality ratings for the same items are 82.02\%, 55.06\%, and 60.67\%. Section 2 (Related Work): ``Our work complements this effort by empirically assessing the current state of prompt reporting in SE research and capturing the expectations of PC members.''
{} additionally derive prompt reporting guidance for automated SE from a review of nearly 300 ICSE, FSE, and ASE papers published since 2022 and a survey of 105 program committee members, finding gaps between current practice and PC member expectations on exact version reporting, prompt construction justification, and treatment of prompts as a validity threat.

\citeauthor{DBLP:journals/corr/abs-2601-01954} frame their empirically derived recommendations as complementing ours, and the essential items their PC respondents endorsed line up with what \modelversion, \design, and \limitationsmitigations already require. Several of \citeauthor{DBLP:conf/chi/NavarroSA26}'s author-side considerations parallel ours as well.%
% GROUNDING: Section 5.1 (page 11) subsection headings: \enq{Transparency regarding prompts and configuration}, \enq{Visualizing system architecture and data flow}, \enq{Evaluating LLM components central to claims}, \enq{Disclosing failure modes of LLM components}, \enq{Justifying the choice to use LLMs}.
{} Their \enq{Transparency regarding prompts and configuration} maps to \modelversion and \design, \enq{Visualizing system architecture and data flow} to \design, \enq{Evaluating LLM components central to claims} to \benchmarksmetrics, \enq{Disclosing failure modes of LLM components} to \limitationsmitigations, and \enq{Justifying the choice to use LLMs} to the empirical comparisons \openllm and \humanvalidation ask for plus the resource-versus-benefit justification in \limitationsmitigations. \citeauthor{DBLP:conf/chi/NavarroSA26}%
% GROUNDING: Navarro et al. Section 6.1.2 ``Comparison to other reporting guidelines and suggestions'': ``Pang et al. [31] recommend that authors use open models and disclose all prompts, while a pre-print in SE asks authors to always `include an open LLM as a baseline' and `publish all prompts' that were considered across the development cycle [6]. Our findings show that authors in HCI have reasonable justifications for going against these guidelines, while the clash of communities suggests HCI needs its own standards.''
{} discuss an earlier arXiv version of our work (arXiv:2508.15503v3, September 2025) in their related work section, summarizing two of our recommendations as \enq{always `include an open LLM as a baseline'} and \enq{publish all prompts}. Neither characterization matches that earlier version or the current version of this manuscript. \openllm states that researchers \should (not \must) include an open LLM as a baseline, specifically for studies that \enq{target commercial tools or models}, with two named exceptions: when human participant tasks make a two-model comparison infeasible, and when LLM use is tangential to the study goal. \design requires the complete set of prompts as \supplementarymaterial only when privacy, anonymity, or confidentiality concerns do not prevent complete disclosure; otherwise summaries and representative examples suffice. Both scoping clauses were already present in the version cited by \citeauthor{DBLP:conf/chi/NavarroSA26}. \citeauthor{DBLP:conf/chi/NavarroSA26}'s own position on prompt reporting%
% GROUNDING: Navarro et al. Section 5.1 ``Transparency regarding prompts and configuration'' (page 11): ``Participants in our study expected the transparent reporting of prompts and configuration details that were central to author claims, or were otherwise difficult to engineer, rather than a universal standard of reporting all prompts.'' ``authors should prioritize reporting prompts and technical details that directly affect reader's ability to validate core claims around system or user behavior.''
{} is more permissive than what \citeauthor{DBLP:journals/corr/abs-2601-01954} report for SE. Their HCI participants prefer reporting only prompts \enq{central to author claims, or [...] otherwise difficult to engineer, rather than a universal standard of reporting all prompts}, whereas 68.54\% of \citeauthor{DBLP:journals/corr/abs-2601-01954}%
% GROUNDING: Korn et al. Section 5, Table 5 essential items (followed in papers / PC member essentiality): ``Authors must provide the exact prompts used word-by-word'' (57.69\% / 68.54\%); ``Authors must describe the prompts used and how they are structured'' (66.08\% / 87.64\%).
{}'s 105 SE PC respondents consider providing the exact prompts \enq{word-by-word} essential and 87.64\% consider describing prompts and their structure essential. Peripheral LLM uses such as proof-reading, spell-checking, and translation are out of scope for these guidelines entirely (see ``Focus on Direct Development or Research Support'' below).

On technical evaluation, \citeauthor{DBLP:conf/chi/NavarroSA26}%
% GROUNDING: Navarro et al. Section 5.2 ``Tempering expectations for rigorous technical evaluations'' (reviewer-side): ``Participants noted that expectations for extensive technical evaluation are sometimes misaligned with the goals of HCI research, where contributions frequently center on interaction paradigms, proofs-of-concept, design explorations, or user understanding rather than benchmarking model behavior. They also cautioned that requests for large-scale or comparative LLM evaluations (e.g., across multiple models or baselines) could impose substantial financial costs.'' Section 5.1 ``Evaluating LLM components central to claims'' (author-side): ``One common approach in recently published CHI and UIST papers [...] is to conduct a modest technical evaluation of the LLM component on a dataset of representative inputs (an `eval' [36]). Datasets are typically custom-tailored to the use case and smaller than formal benchmarks. Metrics might be automated, expert ratings, or both.''
{} argue that \enq{expectations for extensive technical evaluation are sometimes misaligned with the goals of HCI research}, specifying that the evaluations they have in mind are \enq{large-scale or comparative LLM evaluations (e.g., across multiple models or baselines)}, which they say \enq{could impose substantial financial costs}, and recommend instead \enq{a modest technical evaluation of the LLM component on a dataset of representative inputs [...] typically custom-tailored to the use case and smaller than formal benchmarks}. Within our taxonomy of \studytypes, only the subset that automatically evaluates LLM performance, primarily \benchmarkingtasks and applicable \newtools work, falls under \benchmarksmetrics; for those study types we ask for established benchmarks, traditional baselines, and inferential statistics. For the remaining study types (\annotators, \judges, \synthesis, \subjects, \llmusage), evaluation expectations are lighter and the applicability matrix marks \benchmarksmetrics accordingly. The financial cost concern \citeauthor{DBLP:conf/chi/NavarroSA26} raise is addressed in our framework through \limitationsmitigations, which lists resource inequality as a reliability and reproducibility threat, asks researchers to justify the LLM's resource consumption against the benefits over traditional approaches, and treats cost accounting (e.g., input and output tokens, service costs, hardware specifications) as a mitigation strategy.

Our guidelines name \paper or \supplementarymaterial as the reporting location for each recommendation (see Reporting Locations below). The \supplementarymaterial is published according to the \emph{ACM SIGSOFT Open Science Policies}~\cite{daniel_graziotin_2024_10796477} for replication and downstream reuse. \citeauthor{DBLP:conf/chi/NavarroSA26}%
% GROUNDING: Navarro et al. Section 4.2.2 ``Authors use appendices, supplementary material, and repositories to offload LLM details'': ``some authors shifted details into supplementary material or open-source repositories, framing this as a more practical and scalable way to provide access. [...] While this practice helped manage page limits and unwanted distraction from core claims, it also raised questions about how much reviewers actually read or verify supplementary content.''
{} instead describe supplementary material as a venue authors \enq{shifted details into} under page limits, raising \enq{questions about how much reviewers actually read or verify supplementary content}. In summary, our guidelines apply to SE empirical research broadly, organize recommendations around the taxonomy of seven study types introduced in \studytypes with explicit per-study-type applicability, and pair each recommendation with a target reporting location in \paper or \supplementarymaterial.

\scopeparagraph{Focus on Text-Based Use Cases}

While multi-modal foundation models that use or generate images, audio, or video may also support SE research and practice, we focus on textual use cases of LLMs (e.g., in natural language or programming languages). Many of our guidelines, particularly those concerning model reporting, prompt documentation, and reproducibility, are likely applicable to multi-modal settings as well, but we leave their explicit validation to future work.

\scopeparagraph{Focus on Direct Development or Research Support}

While researchers may use LLMs for many peripheral tasks (e.g., proof-reading, spell-checking, translation), our guidelines focus on their direct role in empirical research and engineering practice.
For engineers, we focus on the use of LLMs to automate SE tasks, that is, artificial intelligence (AI) for software engineering (AI4SE) (see~\llmsforengineers).
This includes agentic systems that autonomously plan and execute multi-step tasks using LLMs (see~\design).
For researchers, we focus on the use of LLMs to automate empirical research tasks such as data collection, processing, or analysis (see~\llmsforresearcher).

\scopeparagraph{Researchers as our Target Audience}

Our guidelines are intended to help SE researchers design, plan, conduct,
and report empirical studies involving LLMs, and to support scholarly peer
review of such studies. Each guideline includes an \emph{Advice for
Reviewers} subsection with targeted assessment suggestions.
Our guidelines focus on what to report and how; they complement but do not replace methodological guidance for designing specific types of empirical studies.

\scopeparagraph{Reporting Locations}

We use \paper to refer to the manuscript PDF, including any appendices bound with it. We use \supplementarymaterial to refer to any artifact, replication package, dataset, or other resource external to the manuscript PDF (e.g., hosted on Zenodo or Figshare). What belongs in the main body versus an appendix is a presentation choice for authors and venues; what belongs inside the manuscript versus outside it is the reporting-location decision the guidelines make. When a recommendation does not specify a location, either \paper or \supplementarymaterial is acceptable.

\scopeparagraph{Navigating These Guidelines}

These guidelines are structured to support different reading strategies depending on the reader's goal. \emph{Researchers planning a new study} may start with the taxonomy of study types (see \studytypes) to identify which types apply to their planned work, then consult \ifpaper Table~\ref{tab:guideline-matrix} \else the \href{/guidelines/\#guidelines-by-study-type}{applicability matrix} \fi to determine which guidelines are requirements (\must) and which are recommendations (\should) for those study types. Each guideline section opens with a brief summary in a shaded box, allowing readers to quickly assess relevance before reading the full text. \emph{Researchers writing up results} may prefer to start with the \ifpaper checklist in Appendix~\ref{sec:checklist}, \else \href{/checklist/}{checklist}, \fi which organizes actionable items by typical paper sections (Introduction, Research Design and Methods, Results, etc.). \ifpaper Table~\ref{tab:guidelines} provides a quick reference to all eight guidelines, and Table~\ref{tab:rationale-recommendations} maps each guideline's rationale to its recommendations. \fi \emph{Reviewers} can use the \emph{Advice for Reviewers} subsection at the end of each guideline for targeted guidance on assessing manuscripts. \ifpaper Table~\ref{tab:guideline-matrix} helps reviewers identify which guidelines apply to the study type under review. \fi